Generative adversarial neural networks model of photonic crystal fiber based Surface Plasmon Resonance Sensor

### 2020-10-15 docs/paper/paper.tex
fix:: styling-parameter vals-micro_meter
fixes::by Ahmet

In the last decade, PCF based structures are extensively employed as sensing devices. Possibility of filling the cladding holes with gas and liquid analytes makes them favorable in both chemical and biological sensing applications  \cite{wang2011selectively,rifat2016highly,caucheteur2015review,dash2014spr}. Moreover, due to adaptable light guidance with possibilities of controlling the confinement loss and the chromatic dispersion at any operating wavelength are distinguishing features of PCF based structures in the literature. These charming propagation features can be achieved by employing, circular, cylindrical and even square and rectangular shaped \cite{wang2015simulation} air holes in the cladding and in the core region of the structures. Furthermore, the cladding pattern arrangements such as hexagonal, octagonal, decagonal, square and circular \cite{wang2011selectively,bouk2004dispersion,ademgil2014highly,razzak2007chromatic} can be engaged to achieve the desired features for preferred applications. The hole-to hole spacing (pitch) and the air hole sizes are critical parameters of tuning the aforementioned propagation characteristics. On the other hand, surface plasmon resonance (SPR) is a unique technique that occurs with light and free electrons interaction at metallic boundaries. This interaction results with excitement of surface plasmon polarition (SPP). Relative permittivity variations of the analytes will change the characteristics of the SPP wave, where sensing is triggered. Kretschamann Raether prism geometry is the fundamental construction principle of SPR based sensors  \cite{kretschmann1968radiative}. This geometry practices Snell's Total Internal Reflection rule, where still a very small amount of light refracts from the metallic layer into the evanescent field. The refracted photons are coupling with electron charge oscillations (plasmon) at the surface of the metallic layer, which are known as surface plasmon waves (SPW). The maximum energy transfer from core mode to SPWs occurs with the effective index equalization of these corresponding modes. This phenomenon is known as resonance phase matching, where losses are reaching the maximum levels. The possible changes in the analyte index will directly affect the resonance matching wavelengths. These possible variations are the main parameters for sensitivity calculations. The deep research on SPR based sensors have shown that Kretschmann Raether's prism geometry fails to satisfy the remote sensing and miniaturization requirements of critical detecting applications. One of the most common solutions of overcoming these limitations is replacing the prism with the core of an optical fiber. Considering the recent technological developments, PCF are offering a tremendous amount of design flexibility with the possibilities of complete control of light guidance. Hence, SPR technique can be integrated to this unique microstructured fibers for developing comprehensive multi-purpose sensor \cite{rifat2016highly,yasli2019effect}. Following the exclusive study by Hassani and Skorobogatiy \cite{hassani2006design}, a number of structures have reported the merits of PCF based SPR sensors. In the past, structures that contains features such as multi-channel single analyte and multi-channel multi analyte have been investigated \cite{rifat2016highly,yasli2019effect,zhang2011wagon,yasli2019multi,azzam2016multichannel}. Also, the effects of various metallic layers such as Au, Ag \cite{yasli2019multi}, Ag- Graphene \cite{rifat2016highly}, Ta2TO5 \cite{otupiri2015multi}, ITO \cite{dash2014spr}, TiO2 \cite{rifat2016highly} with various design parameters have also been investigated in detailed. Several design parameters such as plasmonic material, thickness of the layers, air hole size and positions are decisive on resonance matching wavelength of the SPR based structures. Optimizing and refinement of these parameters is an extremely time consuming process with considerable computational time \cite{fornarelli2009neural,abdelaziz2012photonic,paper0}. In the past, genetic algorithms \cite{abdelaziz2012photonic,borel2004topology} method was used to optimize the structural design parameters to achieve better propagation features of photonic devices \cite{rifat2016highly,caucheteur2015review}. Nowadays, in order to optimize such structures, artificial neural networks (ANN) and other machine learning (ML) algorithms are being applied \cite{fornarelli2009neural,paper0,hameed2008accurate}. Although ANNs and ML have been applied to PCFs \cite{fornarelli2009neural,paper0} and SPR structures \cite{fu2018optimization,mcatee2019artificial} independently, to the best of our knowledge, there seems to have been no published papers that investigated the PCF based SPR structures with this well tested methods.

@@

One of the most difficult challenges that deep learning models are facing is that they benefit from large amounts of data to train, which is often costly and/or difficult to acquire. An alternative solution is to artificially expand the original training dataset by the means of generative networks. In this regard, Generative Adversarial Networks (GAN) introduced by Goodfellow et al  \cite{goodfellow2014generative}, has proven to be successful in data generation  \cite{schlegl2017unsupervised, zheng2017unlabeled, frid2018synthetic, tanaka2019data, perez2017effectiveness}.

@@

The GAN can be trained by diverse metrics \cite{goodfellow2014generative, mao2017least, lucic2018gans}]. In this work Wasserstein distance metric (WGAN) is employed \cite{arjovsky2017wasserstein}. In this system, discriminator is named as critic and it measures the distance between the generated and the real data. The distinguishing feature of the WGAN compared to other methods is the ability to determine a stopping criteria. In a regular GAN system, training is stopped when the generated data is deemed viable by an observer. GAN is frequently used in image, video or audio generation, where human is used effectively in the loop. In our case, it is not viable for a human to judge the generated data, therefore, automating this procedure to remove the human in-the-loop can lead to over or under-fitting, which may degrade the performance. On the other hand, WGAN uses an adaptive stopping criteria that can easily overcome the aforementioned disadvantages. Additionally, Gradient penalty is incorporated to improve the WGAN as described in \cite{gulrajani2017improved}. This proposed enhanced WGAN system converges in a stable manner without requiring further fine tuning the hyper-parameters of the system. % The flowchart of this proposed system is illustrated in Figure \ref{fig:gan}.

@@

In order to build the desired dataset, the numerical analysis of the proposed sensor structure is carried out with the commercial software package Comsol Multiphysics  \cite{comsol_help}. The Full Vectorial Finite Element Method (FV-FEM) with the perfectly matched layers (PMLs) \cite{koshiba2002full,saitoh2001full} are employed.

@@

		\end{axis}
		\end{tikzpicture}
		\captionsetup{justification=centering}
		\caption{Configuration estimation on SPR dataset.
		\\\hl{Left: $ n_{analyte} = 1.34 $ $ n_{eff} = 1.46$ $ \Lambda = 0.24 $ $ d1 = 0.45 um $ $ d2= 0.75 um$ $ d3 = 0.35 um $.}\\ \hl{Right: $ n_{analyte} = 1.35 $	$ n_{eff} = 1.45 $ $ \Lambda = 0.15 um $ $ d1 = 0.25 um $ $ d2= 0.75um $ $ d3 = 0.35um $.}		
		  }
	\end{subfigure}
	\caption{Testing results without GAN augmentation phase}
	\label{chart:perfnoaug}
\end{figure*}

@@

		\addlegendimage{first}\addlegendentry{Without GAN:0.32373}
		\addlegendimage{second}\addlegendentry{With GAN:0.08664 }
		\end{axis}
		\end{tikzpicture}
		\captionsetup{justification=centering}
		\caption{Configuration estimation on SPR dataset. 
		\\ \hl{Left: $n_{analyte} = 1.34$ $ n_{eff} = 1.46 $ $ \Lambda = 0.24 um $ $ d1 = 0.45 um $ $ d2= 0.75um $ $ d3 = 0.35um $.} \\ \hl{Right: $n_{analyte} = 1.33 $ $ n_{eff} = 1.45 $ $ \Lambda = 0.15 um $ $ d1 = 0.25 um $ $ d2= 0.55um $ $ d3 = 0.15um $.} 
		}
	\end{subfigure}
	\caption{GAN Phase experimental results}
	\label{chart:ganperf}
\end{figure*}

>>>>

In the last decade, PCF based structures are extensively employed as sensing devices. Possibility of filling the cladding holes with gas and liquid analytes makes them favorable in both chemical and biological sensing applications  \cite{wang2011selectively,rifat2016highly,dash2014spr}. Moreover, due to adaptable light guidance with possibilities of controlling the confinement loss and the chromatic dispersion at any operating wavelength are distinguishing features of PCF based structures in the literature. These charming propagation features can be achieved by employing, circular, cylindrical and even square and rectangular shaped   air holes in the cladding and in the core region of the structures. Furthermore, the cladding pattern arrangements such as hexagonal, octagonal, decagonal, square and circular \cite{wang2011selectively,ademgil2014highly} can be engaged to achieve the desired features for preferred applications. The hole-to hole spacing (pitch) and the air hole sizes are critical parameters of tuning the aforementioned propagation characteristics. On the other hand, surface plasmon resonance (SPR) is a unique technique that occurs with light and free electrons interaction at metallic boundaries. This interaction results with excitement of surface plasmon polarition (SPP). Relative permittivity variations of the analytes will change the characteristics of the SPP wave, where sensing is triggered. Kretschamann Raether prism geometry is the fundamental construction principle of SPR based sensors  \cite{kretschmann1968radiative}. This geometry practices Snell's Total Internal Reflection rule, where still a very small amount of light refracts from the metallic layer into the evanescent field. The refracted photons are coupling with electron charge oscillations (plasmon) at the surface of the metallic layer, which are known as surface plasmon waves (SPW). The maximum energy transfer from core mode to SPWs occurs with the effective index equalization of these corresponding modes. This phenomenon is known as resonance phase matching, where losses are reaching the maximum levels. The possible changes in the analyte index will directly affect the resonance matching wavelengths. These possible variations are the main parameters for sensitivity calculations. The deep research on SPR based sensors have shown that Kretschmann Raether's prism geometry fails to satisfy the remote sensing and miniaturization requirements of critical detecting applications. One of the most common solutions of overcoming these limitations is replacing the prism with the core of an optical fiber. Considering the recent technological developments, PCF are offering a tremendous amount of design flexibility with the possibilities of complete control of light guidance. Hence, SPR technique can be integrated to this unique microstructured fibers for developing comprehensive multi-purpose sensor \cite{rifat2016highly,yasli2019effect}. Following the exclusive study by Hassani and Skorobogatiy \cite{hassani2006design}, a number of structures have reported the merits of PCF based SPR sensors. In the past, structures that contains features such as multi-channel single analyte and multi-channel multi analyte have been investigated \cite{rifat2016highly,yasli2019effect,yasli2019multi,azzam2016multichannel}. Also, the effects of various metallic layers such as Au, Ag \cite{yasli2019multi}, Ag- Graphene \cite{rifat2016highly}, Ta2TO5 \cite{otupiri2015multi}, ITO \cite{dash2014spr}, TiO2 \cite{rifat2016highly} with various design parameters have also been investigated in detailed. Several design parameters such as plasmonic material, thickness of the layers, air hole size and positions are decisive on resonance matching wavelength of the SPR based structures. Optimizing and refinement of these parameters is an extremely time consuming process with considerable computational time \cite{fornarelli2009neural,abdelaziz2012photonic,paper0}. In the past, genetic algorithms \cite{abdelaziz2012photonic,borel2004topology} method was used to optimize the structural design parameters to achieve better propagation features of photonic devices \cite{rifat2016highly}. Nowadays, in order to optimize such structures, artificial neural networks (ANN) and other machine learning (ML) algorithms are being applied \cite{fornarelli2009neural,paper0,hameed2008accurate}. Although ANNs and ML have been applied to PCFs \cite{fornarelli2009neural,paper0} and SPR structures \cite{fu2018optimization,mcatee2019artificial} independently, to the best of our knowledge, there seems to have been no published papers that investigated the PCF based SPR structures with this well tested methods.

@@

One of the most difficult challenges that deep learning models are facing is that they benefit from large amounts of data to train, which is often costly and/or difficult to acquire. An alternative solution is to artificially expand the original training dataset by the means of generative networks. In this regard, Generative Adversarial Networks (GAN) introduced by Goodfellow et al  \cite{goodfellow2014generative}, has proven to be successful in data generation  \cite{schlegl2017unsupervised,  frid2018synthetic,  perez2017effectiveness}.

@@

The GAN can be trained by diverse metrics \cite{goodfellow2014generative, lucic2018gans}]. In this work Wasserstein distance metric (WGAN) is employed \cite{arjovsky2017wasserstein}. In this system, discriminator is named as critic and it measures the distance between the generated and the real data. The distinguishing feature of the WGAN compared to other methods is the ability to determine a stopping criteria. In a regular GAN system, training is stopped when the generated data is deemed viable by an observer. GAN is frequently used in image, video or audio generation, where human is used effectively in the loop. In our case, it is not viable for a human to judge the generated data, therefore, automating this procedure to remove the human in-the-loop can lead to over or under-fitting, which may degrade the performance. On the other hand, WGAN uses an adaptive stopping criteria that can easily overcome the aforementioned disadvantages. Additionally, Gradient penalty is incorporated to improve the WGAN as described in \cite{gulrajani2017improved}. This proposed enhanced WGAN system converges in a stable manner without requiring further fine tuning the hyper-parameters of the system. % The flowchart of this proposed system is illustrated in Figure \ref{fig:gan}.

@@

In order to build the desired dataset, the numerical analysis of the proposed sensor structure is carried out with the commercial software package Comsol Multiphysics. The Full Vectorial Finite Element Method (FV-FEM) with the perfectly matched layers (PMLs) \cite{koshiba2002full} are employed.

@@

		\end{axis}
		\end{tikzpicture}
		\captionsetup{justification=centering}
		\caption{Configuration estimation on SPR dataset.
		\\\hl{Left: $ n_{analyte} = 1.34 $ $ n_{eff} = 1.46$ $ \Lambda = 0.24 \mu m$ $ d_1 = 0.45 \mu m $ $ d_2= 0.75 \mu m$ $ d_3 = 0.35 \mu m $.}\\ \hl{Right: $ n_{analyte} = 1.35 $	$ n_{eff} = 1.45 $ $ \Lambda = 0.15 \mu m $ $ d_1 = 0.25 \mu m $ $ d_2= 0.75 \mu m $ $ d_3 = 0.35 \mu m  $.}		
		  }
	\end{subfigure}
	\caption{Testing results without GAN augmentation phase}
	\label{chart:perfnoaug}
\end{figure*}

@@

		\addlegendimage{first}\addlegendentry{Without GAN:0.32373}
		\addlegendimage{second}\addlegendentry{With GAN:0.08664 }
		\end{axis}
		\end{tikzpicture}
		\captionsetup{justification=centering}
		\caption{Configuration estimation on SPR dataset. 
		\\ \hl{Left: $n_{analyte} = 1.34$ $ n_{eff} = 1.46 $ $ \Lambda = 0.24 \mu m $ $ d_1 = 0.45 \mu m $ $ d_2= 0.75 \mu m $ $ d_3 = 0.35 \mu m $.} \\ \hl{Right: $n_{analyte} = 1.33 $ $ n_{eff} = 1.45 $ $ \Lambda = 0.15 \mu m $ $ d_1 = 0.25 \mu m $ $ d_2= 0.55 \mu m $ $ d_3 = 0.15 \mu m $.} 
		}
	\end{subfigure}
	\caption{GAN Phase experimental results}
	\label{chart:ganperf}
\end{figure*}
